---
title: "How to install packages in R ?"
output: html_document
---

### Option 1: Using Code

#### Install the package
 - `install.packages("package_name")`

 - package name should be in quotes (single or double)

For Example:
```{r eval=FALSE, include=TRUE}
#Install ggplot2 package
install.packages("ggplot2")

#Install tidyr package
install.packages("tidyr")
```

#### Load the package

 - Once the packages are installed, we need to load the packages using the `library(package_name)` command
 - while loading package name should not be in quote

```{r eval=FALSE, include=TRUE}
#Load ggplot2 package
library(ggplot2)

#Load tidyr package
library(tidyr)
```

### Option 2 : Using UI in RStudio/R

#### Install the package

1. Go to Pane : `Files, Plots, Packages, Help, Viewer` Pane which is at Bottom Right in Rstudio
2. Click on Tab : `Packages`
3. Click on `Install`
4. Type the package name which you would like to install

For Example:

 - try installing package named `leaflet`
 - try installing package name `NHANES`
 
#### Load the package

1. Go to Pane : `Files, Plots, Packages, Help, Viewer` Pane which is at Bottom Right in Rstudio
2. Click on Tab : `Packages`
3. Either scroll to look for Package Name or type the package name in search bar
4. Tick/Check on the Check Box (This load the package)

For Example:

 - try loading package  `leaflet`
 - try loading name `NHANES`
